\hypertarget{namespacepyp__graphics}{
\section{Package pyp\_\-graphics}
\label{namespacepyp__graphics}\index{pyp\_\-graphics@{pyp\_\-graphics}}
}


NAME: \hyperlink{pyp__graphics_8py_source}{pyp\_\-graphics.py} VERSION: 2.0.0 (29SEPTEMBER2010) AUTHOR: John B.  




\subsection{Detailed Description}
NAME: \hyperlink{pyp__graphics_8py_source}{pyp\_\-graphics.py} VERSION: 2.0.0 (29SEPTEMBER2010) AUTHOR: John B. Cole, PhD (\href{mailto:john.cole@ars.usda.gov}{\tt john.cole@ars.usda.gov}) LICENSE: LGPL

FUNCTIONS: rmuller\_\-spy\_\-matrix\_\-pil() \mbox{[}1\mbox{]} rmuller\_\-pcolor\_\-matrix\_\-pil() \mbox{[}1\mbox{]} rmuller\_\-get\_\-color() \mbox{[}1\mbox{]} draw\_\-pedigree() plot\_\-founders\_\-by\_\-year() plot\_\-founders\_\-pct\_\-by\_\-year() plot\_\-line\_\-xy() pcolor\_\-matrix\_\-pylab() spy\_\-matrix\_\-pylab() new\_\-draw\_\-pedigree()

\mbox{[}1\mbox{]} These routines were taken from the ASPN Python Cookbook (\href{http://aspn.activestate.com/ASPN/Cookbook/Python/}{\tt http://aspn.activestate.com/ASPN/Cookbook/Python/}) and are used under terms of the license, \char`\"{}Python Cookbook code is freely available for use
     and review\char`\"{} as I understand it. I did not write them; Rick Muller properly deserves credit for that. I THINK Rick's homepage's is: \href{http://www.cs.sandia.gov/~rmuller/.}{\tt http://www.cs.sandia.gov/$\sim$rmuller/.} Python Cookbook notes: Title: Matlab-\/like 'spy' and 'pcolor' functions Submitter: Rick Muller (other recipes) Last Updated: 2005/03/02 Version no: 1.0 Category: Image Description: I really like the 'spy' and 'pcolor' functions, which are useful in viewing matrices. 'spy' prints colored blocks for values that are above a threshold, and 'pcolor' prints out each element in a continuous range of colors. The attached is a little Python/PIL script that does these functions for Numpy arrays. URL: \href{http://aspn.activestate.com/ASPN/Cookbook/Python/Recipe/390208}{\tt http://aspn.activestate.com/ASPN/Cookbook/Python/Recipe/390208} \hyperlink{namespacepyp__graphics}{pyp\_\-graphics} contains routines for working with graphics in PyPedal, such as creating directed graphs from pedigrees using PyDot and visualizing relationship matrices using Rick Muller's spy and pcolor routines (\href{http://aspn.activestate.com/ASPN/Cookbook/Python/}{\tt http://aspn.activestate.com/ASPN/Cookbook/Python/}).

The Python Imaging Library (\href{http://www.pythonware.com/products/pil/}{\tt http://www.pythonware.com/products/pil/}), matplotlib (\href{http://matplotlib.sourceforge.net/}{\tt http://matplotlib.sourceforge.net/}), Graphviz (\href{http://www.graphviz.org/}{\tt http://www.graphviz.org/}), and pydot (\href{http://dkbza.org/pydot.html}{\tt http://dkbza.org/pydot.html}) are required by one or more routines in this module. They ARE NOT distributed with PyPedal and must be installed by the end-\/user! Note that the matplotlib functionality in PyPedal requires only the Agg backend, which means that you do not have to install GTK/PyGTK or WxWidgets/PyWxWidgets just to use PyPedal. Please consult the sites above for licensing and installation information. 