\section*{Legal Notice}
\begin{quote}
 The only people who have anything to fear from free software are those whose products are worth even less. --- David Emery
\end{quote}

\label{sec:legal-notice}
Copyright (c) 2002-2012.  John B. Cole.  All rights reserved.

Permission to use, copy, modify, and distribute this software for any purpose
without fee is hereby granted, provided that this entire notice is included in
all copies of any software which is or includes a copy or modification of this
software and in all copies of the supporting documentation for such software.

\subsection*{Disclaimer}
The author of this software does not make any warranty, express or implied, or
assume any liability or responsibility for the accuracy, completeness, or
usefulness of any information, apparatus, product, or process disclosed, or
represent that its use would not infringe privately-owned rights. Reference
herein to any specific commercial products, process, or service by trade name,
trademark, manufacturer, or otherwise, does not necessarily constitute or imply
its endorsement, recommendation, or favoring by the United States Government or
the author. The views and opinions of authors expressed herein do not necessarily
state or reflect those of the United States Government and shall not be used for
advertising or product endorsement purposes.

\section*{License}\label{sec:license}\index{license}
This is free software; you can redistribute it and/or modify it
under the terms of the GNU Lesser General Public License as published by
the Free Software Foundation; either version 2.1 of the License, or (at
your option) any later version.

This software is distributed in the hope that it will be useful, but
WITHOUT ANY WARRANTY; without even the implied warranty of
MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
Lesser General Public License for more details.

You should have received a copy of the GNU Lesser General Public
License along with this library; if not, write to the Free Software
Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301,
USA.